---
title: "LDA"
author: "Jason"
date: "June 20, 2016"
output: pdf_document
---

**7. LDA - 22 Second Bin Width (1100 time points)**

```{r}
#DCM:
library(MASS)
lda_info = lda(as.factor(dataframe[,3])~binned_matrix_1100)
lda_info$scaling #Linear discriminant coefficient matrix
lda_info_p = predict(lda_info,as.data.frame(binned_matrix_1100))
lda_info_pclass = predict(lda_info)$class
par(mar=c(5,4,4,4))
v = as.numeric(lda_info_pclass)
v[which(v==1)] =  "deeppink"   #D
v[which(v==2)] =  12           #F
v[which(v==3)] =  "grey"       #Trochi

plot(lda_info_p$x[,1],pch=16, col = v,main = "LDA - 22 Second Bin Width",xlab="df1",ylab="df2") #Plot the linear discriminants
legend("topright",legend=c("S. glaucum clade D","S. glaucum clade F","S. trocheliophorum"),col=c("deeppink",12,"grey"), pch=16,bty='o')

clade.manova = manova(binned_matrix_1100~as.factor(dataframe[,3]))
clade.wilks = summary(clade.manova,test="Wilks");clade.wilks

#HEX:
library(MASS)
lda_info_HEX = lda(as.factor(dataframeHEX[,3])~binned_matrixHEX_1100)
lda_info_HEX$scaling #Linear discriminant coefficient matrix
lda_info_p = predict(lda_info_HEX,as.data.frame(binned_matrixHEX_1100))
lda_info_pclass = predict(lda_info_HEX)$class
par(mar=c(5,4,4,4))
v = as.numeric(lda_info_pclass)
v[which(v==1)] =  "deeppink"   #D
v[which(v==2)] =  12           #F
v[which(v==3)] =  "grey"       #Trochi

plot(lda_info_p$x[,1],pch=16, col = v,main = "LDA - 22 Second Bin Width",xlab="df1",ylab="df2") #Plot the linear discriminants
legend("topright",legend=c("S. glaucum clade D","S. glaucum clade F","S. trocheliophorum"),col=c("deeppink",12,"grey"), pch=16,bty='o')

clade.manova = manova(binned_matrix_1100~as.factor(dataframe[,3]))
clade.wilks = summary(clade.manova,test="Wilks");clade.wilks

```

The Wilk's Lambda statistic corresponds to a p-value of ~.09, suggesting that the clades account for a large (but only marginally significant) proportion of the variance in binned area.

**8. LDA - 40 Second Bin Width (2000 time points)**

```{r}
#DCM:
lda_info = lda(as.factor(dataframe[,3])~binned_matrix_2000)
lda_info$scaling  #Linear discriminant coefficient matrix
lda_info_p = predict(lda_info,as.data.frame(binned_matrix_2000))
lda_info_pclass = predict(lda_info)$class
par(mar=c(5,4,4,4))
v = as.numeric(lda_info_pclass)
v[which(v==1)] =  "deeppink"   #D
v[which(v==2)] =  12           #F
v[which(v==3)] =  "grey"       #Trochi

plot(lda_info_p$x[,1],pch=16, col = v,main = "LDA - 40 Second Bin Width",xlab="df1",ylab="df2") #Plot the linear discriminants
legend("topright",legend=c("S. glaucum clade D","S. glaucum clade F","S. trocheliophorum"),col=c("deeppink",12,"grey"), pch=16,bty='o')

clade.manova = manova(binned_matrix_2000~as.factor(dataframe[,3]))
clade.wilks = summary(clade.manova,test="Wilks");clade.wilks

#HEX:
library(MASS)
lda_info_HEX = lda(as.factor(dataframeHEX[,3])~binned_matrixHEX_2000)
lda_info_HEX$scaling #Linear discriminant coefficient matrix
lda_info_p = predict(lda_info_HEX,as.data.frame(binned_matrixHEX_2000))
lda_info_pclass = predict(lda_info_HEX)$class
par(mar=c(5,4,4,4))
v = as.numeric(lda_info_pclass)
v[which(v==1)] =  "deeppink"   #D
v[which(v==2)] =  12           #F
v[which(v==3)] =  "grey"       #Trochi

plot(lda_info_p$x[,1],pch=16, col = v,main = "LDA - 22 Second Bin Width",xlab="df1",ylab="df2") #Plot the linear discriminants
legend("topright",legend=c("S. glaucum clade D","S. glaucum clade F","S. trocheliophorum"),col=c("deeppink",12,"grey"), pch=16,bty='o')

clade.manova = manova(binned_matrix_2000~as.factor(dataframe[,3]))
clade.wilks = summary(clade.manova,test="Wilks");clade.wilks
```

The Wilk's Lambda statistic corresponds to a p-value of ~0.0008, suggesting that the clades account for a large, significant proportion of the variance in binned area.

**9. LDA — 60 Second Bin Width (3000 time points)**

```{r}
#DCM:
lda_info = lda(as.factor(dataframe[,3])~binned_matrix_3000)
lda_info$scaling  #Linear discriminant coefficient matrix
lda_info_p = predict(lda_info,as.data.frame(binned_matrix_3000))
lda_info_pclass = predict(lda_info)$class
par(mar=c(5,4,4,4))
v = as.numeric(lda_info_pclass)
v[which(v==1)] =  "deeppink"   #D
v[which(v==2)] =  12           #F
v[which(v==3)] =  "grey"       #Trochi

plot(lda_info_p$x[,1],pch=16, col = v,main = "LDA - 60 Second Bin Width",xlab="df1",ylab="df2") #Plot the linear discriminants
legend("topright",legend=c("S. glaucum clade D","S. glaucum clade F","S. trocheliophorum"),col=c("deeppink",12,"grey"), pch=16,bty='o')

clade.manova = manova(binned_matrix_3000~as.factor(dataframe[,3]))
clade.wilks = summary(clade.manova,test="Wilks");clade.wilks

#HEX:
library(MASS)
lda_info_HEX = lda(as.factor(dataframeHEX[,3])~binned_matrixHEX_3000)
lda_info_HEX$scaling #Linear discriminant coefficient matrix
lda_info_p = predict(lda_info_HEX,as.data.frame(binned_matrixHEX_3000))
lda_info_pclass = predict(lda_info_HEX)$class
par(mar=c(5,4,4,4))
v = as.numeric(lda_info_pclass)
v[which(v==1)] =  "deeppink"   #D
v[which(v==2)] =  12           #F
v[which(v==3)] =  "grey"       #Trochi

plot(lda_info_p$x[,1],pch=16, col = v,main = "LDA - 22 Second Bin Width",xlab="df1",ylab="df2") #Plot the linear discriminants
legend("topright",legend=c("S. glaucum clade D","S. glaucum clade F","S. trocheliophorum"),col=c("deeppink",12,"grey"), pch=16,bty='o')

clade.manova = manova(binned_matrix_3000~as.factor(dataframe[,3]))
clade.wilks = summary(clade.manova,test="Wilks");clade.wilks

```

The Wilk's Lambda statistic corresponds to a p-value of ~0, suggesting that the clades account for a large, significant proportion of the variance in binned area.
\

